---
title: "Decorate Module Output"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Decorate Module Output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(teal.osprey)
```


## Introduction

The outputs produced by `teal` modules, like graphs or tables, are created by the module developer and look a certain way.
It is hard to design an output that will satisfy every possible user, so the form of the output should be considered a default value that can be customized.
Here we describe the concept of _decoration_, enabling the app developer to tailor outputs to their specific requirements without rewriting the original module code.

The decoration process is build upon transformation procedures, introduced in `teal`.
While `transformators` are meant to edit module's input, decorators are meant to adjust the module's output.
To distinguish the difference, modules in  `teal.osprey` have 2 separate parameters:
`transformators` and `decorators`. 

To get a complete understanding refer the following vignettes:

- Transforming the input data in
[this vignette](https://insightsengineering.github.io/teal/latest-tag/articles/transform-input-data.html).
- Transforming module output in [this vignette](https://insightsengineering.github.io/teal/latest-tag/articles/transform-module-output.html).


## Outputs that can be decorated

It is important to note which output objects from a given module can be decorated.
The module function documentation's _Decorating Module_ section has this information.

You can also refer the table shown below to know which module outputs can be decorated.


| Module                  | Output  (Class) |
|-------------------------|----------------------------|
| `tm_g_spiderplot`       | plot  `ggplot` |
| `tm_g_butterfly`        | plot  `grob` / `gtable` |
| `tm_g_waterfall`        | plot  `grob` / `gtable` |
| `tm_g_swimlane`         | plot  `grob` / `gtable` |
| `tm_g_patient_profile`  | plot  `grob` (`cowplot` layout) |
| `tm_g_ae_oview`         | plot  `grob` |
| `tm_g_ae_sub`           | plot  `grob` |
| `tm_g_events_term_id`   | plot  `grob` |
| `tm_g_heat_bygrade`     | plot  `grob` / `gtable` |



Also, note that there are four different types of objects that can be decorated:

1. `ggplot`
2. `grob`
3. `gtable`
4. `cowplot`



## Decorating `ggplot`


Here's an example to showcase how you can edit an output of class `ggplot`.
You can extend them using `ggplot2` functions.

```{r decorate_ggplot, message=FALSE}
library(teal.modules.general)

data <- teal_data(join_keys = default_cdisc_join_keys[c("ADSL", "ADRS")])
data <- within(data, {
  require(nestcolor)
  ADSL <- rADSL
})

ggplot_caption_decorator <- function(default_caption = "I am a good decorator") {
  teal_transform_module(
    label = "Caption",
    ui = function(id) {
      shiny::textInput(shiny::NS(id, "footnote"), "Footnote", value = default_caption)
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          data() |>
            within(
              {
                plot <- plot + ggplot2::labs(caption = footnote)
              },
              footnote = input$footnote
            )
        })
      })
    }
  )
}

app <- init(
  data = data,
  modules = modules(
    tm_g_spiderplot(
      decorators = list(
        plot = ggplot_caption_decorator("I am a Regression")
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_2, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("decorate_ggplot")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```

## Decorating `grob`

Here's an example to showcase how you can edit an output of class `grob`.
You can extend them using `grid` and `gridExtra` functions.

```{r decorate_grob, message=FALSE}
library(teal.modules.general)

data <- teal_data(join_keys = default_cdisc_join_keys[c("ADSL", "ADRS")])
data <- within(data, {
  ADSL <- rADSL
})

grob_caption_decorator <- function(default_caption = "I am a good decorator") {
  teal_transform_module(
    label = "Caption",
    ui = function(id) {
      shiny::textInput(shiny::NS(id, "footnote"), "Footnote", value = default_caption)
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          data() |>
            within(
              {
                footnote_grob <- grid::textGrob(
                  footnote,
                  x = 0, hjust = 0,
                  gp = grid::gpar(fontsize = 10, fontface = "italic", col = "gray50")
                )
                plot <- gridExtra::arrangeGrob(
                  plot,
                  footnote_grob,
                  ncol = 1,
                  heights = grid::unit.c(
                    grid::unit(1, "npc") - grid::unit(1, "lines"), grid::unit(1, "lines")
                  )
                )
              },
              footnote = input$footnote
            )
        })
      })
    }
  )
}

app <- init(
  data = data,
  modules = modules(
    tm_g_ae_sub(
      decorators = list(
        plot = grob_caption_decorator("I am a Association")
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_3, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("decorate_grob")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```

## Decorating `gtable`

## Decorating `cowplot`