---
title: "Decorate Module Output"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Decorate Module Output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(teal.osprey)
```


## Introduction

The outputs produced by `teal` modules, like graphs or tables, are created by the module developer and look a certain way.
It is hard to design an output that will satisfy every possible user, so the form of the output should be considered a default value that can be customized.
Here we describe the concept of _decoration_, enabling the app developer to tailor outputs to their specific requirements without rewriting the original module code.

The decoration process is build upon transformation procedures, introduced in `teal`.
While `transformators` are meant to edit module's input, decorators are meant to adjust the module's output.
To distinguish the difference, modules in  `teal.osprey` have 2 separate parameters:
`transformators` and `decorators`. 

To get a complete understanding refer the following vignettes:

- Transforming the input data in
[this vignette](https://insightsengineering.github.io/teal/latest-tag/articles/transform-input-data.html).
- Transforming module output in [this vignette](https://insightsengineering.github.io/teal/latest-tag/articles/transform-module-output.html).


## Outputs that can be decorated

It is important to note which output objects from a given module can be decorated.
The module function documentation's _Decorating Module_ section has this information.

You can also refer the table shown below to know which module outputs can be decorated.


| Module                  | Output  (Class) |
|-------------------------|----------------------------|
| `tm_g_spiderplot`       | plot  `ggplot` |
| `tm_g_butterfly`        | plot  `grob` / `gtable` |
| `tm_g_waterfall`        | plot  `grob` / `gtable` |
| `tm_g_swimlane`         | plot  `grob` / `gtable` |
| `tm_g_patient_profile`  | plot  `grob` / `ggplot` |
| `tm_g_ae_oview`         | plot  `grob` |
| `tm_g_ae_sub`           | plot  `grob` |
| `tm_g_events_term_id`   | plot  `grob` |
| `tm_g_heat_bygrade`     | plot  `grob` / `gtable` |



Also, note that there are three different types of objects that can be decorated:

1. `ggplot`
2. `grob`
3. `gtable`



## Decorating `ggplot`


Here's an example to showcase how you can edit an output of class `ggplot`.
You can extend them using `ggplot2` functions.

```{r decorate_ggplot, message=FALSE, eval = FALSE}
library(teal.osprey)

data <- within(teal_data(), {
  require(nestcolor)
  ADSL <- rADSL
  ADTR <- rADTR
})

join_keys(data) <- default_cdisc_join_keys[names(data)]

ggplot_caption_decorator <- function(default_caption = "I am a good decorator") {
  teal_transform_module(
    label = "Caption",
    ui = function(id) {
      shiny::textInput(shiny::NS(id, "footnote"), "Footnote", value = default_caption)
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          data() |>
            within(
              {
                plot <- plot + ggplot2::labs(caption = footnote)
              },
              footnote = input$footnote
            )
        })
      })
    }
  )
}

app <- init(
  data = data,
  modules = modules(
    tm_g_spiderplot(
      label = "Spiderplot",
      dataname = "ADTR",
      paramcd = variables(
        choices = is_categorical(),
        selected = "PARAMCD"
      ),
      x_var = variables(
        choices = dplyr::where(is.numeric),
        selected = 1L
      ),
      y_var = variables(
        choices = c("PCHG", "CHG", "AVAL"),
        selected = "PCHG"
      ),
      marker_var = variables(
        choices = c("SEX", "RACE", "USUBJID"),
        selected = "SEX"
      ),
      line_colorby_var = variables(
        choices = c("SEX", "USUBJID", "RACE"),
        selected = "SEX"
      ),
      xfacet_var = variables(
        choices = c("SEX", "ARM"),
        selected = "SEX"
      ),
      yfacet_var = variables(
        choices = c("SEX", "ARM"),
        selected = "ARM"
      ),
      decorators = list(
        plot = ggplot_caption_decorator("I am a ggplot")
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_1, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("decorate_ggplot")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```

## Decorating `grob`

Here's an example to showcase how you can edit an output of class `grob`.
You can extend them using `grid` and `gridExtra` functions.

```{r decorate_grob, message=FALSE, eval=FALSE}
library(teal.osprey)

data <- within(teal_data(), {
  ADSL <- rADSL
  ADTR <- rADTR
})

join_keys(data) <- default_cdisc_join_keys[names(data)]

grob_caption_decorator <- function(default_caption = "I am a good decorator") {
  teal_transform_module(
    label = "Caption",
    ui = function(id) {
      shiny::textInput(shiny::NS(id, "footnote"), "Footnote", value = default_caption)
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          data() |>
            within(
              {
                footnote_grob <- grid::textGrob(
                  footnote,
                  x = 0, hjust = 0,
                  gp = grid::gpar(fontsize = 10, fontface = "italic", col = "gray50")
                )
                plot <- gridExtra::arrangeGrob(
                  plot,
                  footnote_grob,
                  ncol = 1,
                  heights = grid::unit.c(
                    grid::unit(1, "npc") - grid::unit(1, "lines"), grid::unit(1, "lines")
                  )
                )
              },
              footnote = input$footnote
            )
        })
      })
    }
  )
}

app <- init(
  data = data,
  modules = modules(
    tm_g_ae_sub(
      label = "AE subview",
      dataname = "ADTR",
      arm_var = variables(
        choices = is_categorical(),
        selected = "ACTARMCD"
      ),
      group_var = variables(
        choices = is_categorical(),
        selected = c("SEX", "REGION1", "RACE")
      ),
      decorators = list(
        plot = grob_caption_decorator("I am a grob")
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_2, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("decorate_grob")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```

## Decorating `gtable`


Here's an example to showcase how you can edit an output of class `gtable`.
You can extend them using `grid` and `gridExtra` functions.

```{r decorate_gtable, message=FALSE, eval=FALSE}
library(teal.osprey)
library("dplyr")
data <- within(teal_data(), {
  library("dplyr")
  ADSL <- rADSL %>%
    mutate(TRTDURD = as.integer(TRTEDTM - TRTSDTM) + 1) %>%
    filter(STRATA1 == "A" & ARMCD == "ARM A")
  ADRS <- rADRS %>%
    filter(PARAMCD == "LSTASDI" & DCSREAS == "Death") %>%
    mutate(AVALC = DCSREAS, ADY = EOSDY) %>%
    rbind(rADRS %>% filter(PARAMCD == "OVRINV" & AVALC != "NE")) %>%
    arrange(USUBJID)
})

join_keys(data) <- default_cdisc_join_keys[names(data)]

gtable_caption_decorator <- function(default_caption = "I am a good decorator") {
  teal_transform_module(
    label = "Caption",
    ui = function(id) {
      shiny::textInput(shiny::NS(id, "footnote"), "Footnote", value = default_caption)
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          data() |>
            within(
              {
                footnote_grob <- grid::textGrob(
                  footnote,
                  x = 0, hjust = 0,
                  gp = grid::gpar(fontsize = 10, fontface = "italic", col = "gray50")
                )
                plot <- gtable::gtable_add_grob(
                  plot,
                  footnote_grob,
                  t = 1,
                  l = 1,
                  clip = "on"
                )
              },
              footnote = input$footnote
            )
        })
      })
    }
  )
}

app <- init(
  data = data,
  modules = modules(
    tm_g_swimlane(
      label = "Swimlane Plot",
      dataname = "ADRS",
      bar_var = variables(
        choices = c("TRTDURD", "EOSDY"),
        selected = "TRTDURD"
      ),
      bar_color_var = variables(
        choices = c("EOSSTT", "ARM", "ARMCD", "ACTARM", "ACTARMCD", "SEX"),
        selected = "EOSSTT"
      ),
      decorators = list(
        plot = gtable_caption_decorator("I am a gtable")
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_3, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("decorate_gtable")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```

